---
title: "Data Visulaization on Gapminder Dataframe"
author: "Vinodh Reddy Chennu"
date: "Reg n.o: 18MBMB09; MBA Business Analytics"
output:
  
  
  ioslides_presentation: default
  beamer_presentation: default
  slidy_presentation: default
---
<style type="text/css">

body, td {
   font-size: 14px;
}
code.r{
  font-size: 12px;
}
pre {
  font-size: 12px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
```

## Gapminder DataFrame
Gapminder data on life expectancy, GDP per capita and population by country.

>The main data frame 'gapminder' has 1704 rows and 6 variables(columns).

###**Variables includes:**
1. **country**: factor with 142 levels.
2. **continent**: factor with 5 levels.
3. **year**: ranges from 1952 to 2007 in increments of 5 years.
4. **lifeExp**: life expectancy at birth, in years.
5. **pop**: population.
6. **gdpPercap**: GDP per capita (US$, inflation-adjusted).

## Summary of Gapminder Dataframe

```{r gapminder, size="huge",echo=TRUE}
summary(gapminder::gapminder)
```

## Plotted Visualizations
- **Bar Chart** | TotalPop by Continents
- **Box Plot** | LifeExp by continents
- **ScatterPlot** | gdpPercap vs lifeExp
- **ScatterPlot** | gdpPercap vs lifeExp on log scale with continents differentiated with colors & pop with size
- **ScatterPlots with facetwrap** | gdpPercap vs LifeExp of each continent
- **Line graph** | Total population of continents over years
- **Line graph** | Population over the years and continents

##Create dataframe with filter year=2007
```{r dataset 2007, echo=TRUE}
gapminder_year2007 <- gapminder::gapminder %>% filter(year == 2007)
gapminder_year2007
```

##Population by continents

```{r bar chart Mean life by continent , echo=TRUE}
ggplot(gapminder_year2007, aes(x = continent, y = pop)) + geom_col()
```

##LifeExp by Continents

```{r boxplot, echo=TRUE}
ggplot(gapminder_year2007, aes(x = continent, y = lifeExp)) + geom_boxplot()
```

## gdpPercap vs lifeExp

```{r scatter plot pop vs lifexep, echo=TRUE}
ggplot(gapminder_year2007, aes(x = gdpPercap, y = lifeExp)) + geom_point()
```

## gdpPercap vs lifeExp with more options

```{r scatter plot, echo=TRUE}
ggplot(gapminder_year2007, aes(x = gdpPercap, y = lifeExp, color = continent, size=pop)) + geom_point() + scale_x_log10()
```

##gdpPercap vs LifeExp of each continent

```{r facetyear, echo=TRUE}
ggplot(gapminder_year2007, aes(x = gdpPercap, y = lifeExp,color = continent, size=pop)) + geom_point() + scale_x_log10() +
  facet_wrap(~ continent)
```

## Created dataframe with groupby(year)

```{r lineplot2, echo=TRUE}
by_year <- gapminder::gapminder %>% group_by(year) %>% summarize(totalPop = sum(as.numeric(pop)), meanLifeExp = mean(lifeExp))
by_year
```

## Total population over the years

```{r pop over years, echo=TRUE}
ggplot(by_year, aes(x = year, y = totalPop)) + geom_line()
```

##Data with groupby(year,continent)
```{r by_year_continent,echo=TRUE}
by_year_continent <- gapminder::gapminder %>% group_by(year, continent) %>% summarize(totalPop = sum(as.numeric(pop)),
meanLifeExp = mean(lifeExp))
by_year_continent
```

##Population over the year by continents

```{r Visualizing population by year and continent, echo=TRUE}
ggplot(by_year_continent, aes(x = year, y = totalPop, color = continent)) + geom_line() + expand_limits(y = 0)
```

## Conclusions

1. Asia continent has the higher population Whereas Ocenia continent has the least population.
2. Median lifeExp of Africa Continent is least and Ocenia continent is the highest.
3. On gdpPerCap vs lifeExp plots, most of the Africa countries lies at bottom left whereas Americas, Ocieania and European continents lie at the top right. There is a correlation between lifeExp and gdpPercap.
4. Total Population on the World increases over the time.
5. Total Population of Asia, Africa, Americas continents increases rapidly Whereas Europe and Ocenia changes very slightly over the time.
